Beyond the Numbers: Dissecting New York City's Leading Causes of Death Across Demographics

### 2024-03-17 paper/paper.qmd
updated graphs

# Prepare the data
avg_rate_top_causes_by_race <- nyc_data %>%
  mutate(Age_Adjusted_Death_Rate = as.numeric(Age_Adjusted_Death_Rate)) %>%
  group_by(Race.Ethnicity, Leading.Cause) %>%
  summarise(Average_Age_Adjusted_Death_Rate = mean(Age_Adjusted_Death_Rate, na.rm = TRUE), .groups = 'drop') %>%
  arrange(Race.Ethnicity, desc(Average_Age_Adjusted_Death_Rate)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= 5) %>%
  select(Race.Ethnicity, Leading.Cause, Average_Age_Adjusted_Death_Rate)

# Plotting
gg <- ggplot(avg_rate_top_causes_by_race, aes(x = `Leading.Cause`, y = Average_Age_Adjusted_Death_Rate, fill = `Leading.Cause`)) +
  geom_col() + 
  facet_wrap(~`Race.Ethnicity`, scales = "free_y") + 
  scale_fill_brewer(palette = "Paired") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 40, hjust = 1), 
        strip.text.x = element_text(size = 6), 
        legend.position = "none") + 
  labs(x = "Top 5 Causes of Death", y = "Average Age-Adjusted Death Rate")

print(gg)
ggsave("avg_rate_top_causes_by_race.png", plot = gg, width = 12, height = 8)

>>>>

gg <- ggplot(avg_rate_top_causes_by_race, aes(x = Leading.Cause, y = Average_Age_Adjusted_Death_Rate)) +
  geom_col(aes(fill = Leading.Cause)) + 
  facet_wrap(~Race.Ethnicity, scales = "free_y") + 
  scale_fill_brewer(palette = "Paired") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 40, hjust = 1), 
        strip.text.x = element_text(size = 6), 
        legend.position = "none") + 
  labs(x = "Top 5 Causes of Death", y = "Average Age-Adjusted Death Rate")

print(gg)
